---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data

```{r}
library(ggplot2, quietly = TRUE, verbose = FALSE)
library(dplyr, quietly = TRUE, verbose = FALSE, warn.conflicts = FALSE)
library(chron, quietly = TRUE, verbose = FALSE)
dsClasses <- c("numeric", "character", "numeric")
ds <- read.csv("activity.csv", header = TRUE, colClasses = dsClasses, na.strings=c("","NA"))
ds$date <- as.Date(ds$date, format="%Y-%m-%d")
```

## Total number of steps taken per day

Below, you can see the histogram of the mean and median of the total number of steps taken per day.

```{r}
totalStepsByDate = summarise(group_by(ds, date), totalSteps = sum(steps))
qplot(totalStepsByDate$totalSteps, geom="histogram") + xlab("Steps") + ggtitle("Total Steps By Day")
```

## Total mean number of steps taken per day

The mean is `r as.character(mean(totalStepsByDate$totalSteps, na.rm = TRUE))`

The average is `r as.character(median(totalStepsByDate$totalSteps, na.rm = TRUE))`

## The average daily activity pattern

This is a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
averageStepsByInterval = summarise(group_by(ds, interval), averageSteps = mean(steps, na.rm = TRUE))
ggplot(averageStepsByInterval, aes(interval, averageSteps)) + geom_line() + xlab("Interval") + ylab("Average Steps") + ggtitle("Average Steps by Interval")
```

The maximum number of steps (5-minute interval, on average across all the days in the dataset) is `r as.character(filter(averageStepsByInterval, averageSteps == max(averageStepsByInterval$averageSteps)))`

## Imputing missing values

Total number of missing values in the dataset is `r length(ds$steps[is.na(ds$steps) == TRUE])`

You can see that NA values effectively change the results. Compare to analysis below.

```{r}
ds$steps[is.na(ds$steps) == TRUE] <- 0
```

Below is a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.

```{r}

totalStepsByDate = summarise(group_by(ds, date), totalSteps = sum(steps))
qplot(totalStepsByDate$totalSteps, geom="histogram") + xlab("Steps") + ggtitle("Total Steps By Day")

```

The mean is `r as.character(mean(totalStepsByDate$totalSteps, na.rm = TRUE))`

The average is `r as.character(median(totalStepsByDate$totalSteps, na.rm = TRUE))`

## Differences in activity patterns between weekdays and weekends

```{r}
daysOfWeek <- is.weekend(ds$date)
ds$isWeekend <- factor(daysOfWeek, labels = c("Weekday", "Weekend"))
averageStepsByInterval = summarise(group_by(ds, interval, isWeekend), averageSteps = mean(steps))

g <- ggplot(averageStepsByInterval, aes(interval, averageSteps)) + geom_line() + xlab("Interval") + ylab("Average Steps") + ggtitle("Average Steps by Interval")
g + facet_grid(isWeekend ~ .)
```
